---
title: "R markdown demo"
---

Demo of some data wrangling using a subset of the World Bank Development Indicators "Excel download"" from:

http://datatopics.worldbank.org/world-development-indicators/


```{r setup, include=TRUE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, include = TRUE, cache = FALSE, warning = FALSE, message = FALSE)
library(knitr)
library(tidyverse)
```

#### Read in the data as data frames

Data is available at https://github.com/actsci-UoL/DSAdemo/tree/master/Data and needs to be saved locally in a sub-directory of the working directory called "Data". 

Warning: one of the files is over 50Mb.

```{r}
indicators <- 
  read_csv("Data/Indicators.csv") %>% 
  mutate(Country_Code = as.factor(Country_Code),
         Indicator_Code = as.factor(Indicator_Code))
country <- 
  read_csv("Data/Country.csv") %>% 
  mutate(Country_Code = as.factor(Country_Code),
         Region = as.factor(Region),
         Income_Group = as.factor(Income_Group))
```



## Questions

For each question below (not the first), find the answer using the `dplyr` package. In a later tutorial we will run through exctly the same sequence using SQL

  * How many different indicators and countries are there?
  * List the regional groups
  * List the country names and codes
  * Which country codes are not linked to a region?
  * List the indicators that contain 'life expectancy' in the indicator name
  * Find the average, minimum, and maximum female life expectancy at birth in 2016
  * Find the name of the country with the minimum female life expectancy at birth
    +  Print the countries with life expectancy < 60 together with their region and Income group
    +  Print the countries with life expectancy > 85 together with their region and Income group
  * Plot life expectancy at birth vs percentage of births attended by a qualified medic
  * Calculate the average female life expectancy at birth in 2016 for each region and income group.
    +  Rename the life expectancy column to something meaningful and order by region and life expectancy.
    

## Solutions


#### How many different indicators and countries are there?

```{r}
#
```


```{r}
#
```



#### List the regional groups


```{r}
#
```


#### List the country names and codes

```{r}
#
```


#### Which country codes are not linked to a region?

```{r}
#
``` 



#### List the indicators that contain 'life expectancy' in the indicator name


```{r}
#
```



#### Find the average, minimum, and maximum female life expectancy at birth in 2016

```{r}
#
```


#### Find the name of the country with the minimum female life expectancy at birth


```{r}
#
```


#### Plot life expectancy at birth vs percentage of births attended by a qualified medic




```{r}
# Find out the name of the indicator and its code.

```



```{r, include = TRUE}
# Make a plot.

```


#### Use a join to add Region and Income Group to the indicators table



```{r}
# Print the countries with life expectancy > 85 together with their region and Income group.

```




```{r}
# Print the countries with life expectancy < 60 together with their region and Income group.

```


#### Calculate the average female life expectancy at birth in 2017 for each region and income group.



```{r}
# Rename the life expectancy column to something meaningful and order by region and life expectancy.


```


