---
title: "R markdown demo"
---

Demo of some data wrangling using a subset of the World Bank Development Indicators "Excel download"" from:

http://datatopics.worldbank.org/world-development-indicators/


```{r setup, include=TRUE, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, include = FALSE, cache = TRUE, warning = FALSE,
                      message = FALSE, dplyr.summarise.inform = FALSE)
library(knitr)
library(tidyverse)
```

#### Read in the data as data frames

Data is available at https://github.com/actsci-UoL/DSAdemo/tree/master/Data and needs to be saved locally in a sub-directory of the working directory called "Data". 

Warning: one of the files is over 50Mb.

```{r}
indicators <- 
  read_csv("Data/Indicators.csv") %>% 
  mutate(Country_Code = as.factor(Country_Code),
         Indicator_Code = as.factor(Indicator_Code))
country <- 
  read_csv("Data/Country.csv") %>% 
  mutate(Country_Code = as.factor(Country_Code),
         Region = as.factor(Region),
         Income_Group = as.factor(Income_Group))
```



## Questions

For each question below (not the first), find the answer using the `dplyr` package. In a later tutorial we will run through exctly the same sequence using SQL

  * How many different indicators and countries are there?
  * List the regional groups
  * List the country names and codes
  * Which country codes are not linked to a region?
  * List the indicators that contain 'life expectancy' in the indicator name
  * Find the average, minimum, and maximum female life expectancy at birth in 2016
  * Find the name of the country with the minimum female life expectancy at birth
    +  Print the countries with life expectancy < 60 together with their region and Income group
    +  Print the countries with life expectancy > 85 together with their region and Income group
  * Plot life expectancy at birth vs percentage of births attended by a qualified medic
  * Calculate the average female life expectancy at birth in 2016 for each region and income group.
    +  Rename the life expectancy column to something meaningful and order by region and life expectancy.
    

## Model solutions


#### How many different indicators and countries are there?

```{r}
num_indic <-
  indicators %>%
  summarise(Num_indicators = n_distinct(Indicator_Code, na.rm = TRUE))
num_indic
```
There are `r num_indic[[1]]` different indicators.

```{r}
num_coun <- 
  indicators %>%  
  summarise(Num_countries = n_distinct(Country_Code, na.rm = TRUE))
num_coun
```
There are `r num_coun[[1]]` different countries.


#### List the regional groups


```{r}
reg_groups <- 
  country %>% 
  filter(!is.na(Region)) %>% 
  select(Region) %>% 
  unique()
reg_groups
```
The regional groups are: `r kable(reg_groups)`.

#### List the country names and codes

```{r}
country_table <- 
  indicators %>% 
  select(Country_Code, Country_Name) %>% 
  unique() %>% 
  arrange(Country_Code) %>% 
  sample_n(20)
country_table
```
A sample of countries and country codes: `r kable(country_table)`

#### Which country codes are not linked to a region?

```{r}
nr_country_table <- 
  country %>% 
  filter(is.na(Region)) %>% 
  select(Country_Code) %>% 
  unique()
nr_country_table
``` 
Country codes not associated with a region: `r kable(nr_country_table)`


#### List the indicators that contain 'life expectancy' in the indicator name


```{r}
le_indics <- 
  indicators %>% 
  filter(str_detect(Indicator_Name, "(?i)life expectancy")) %>% # (?i) says ignore case
  select(Indicator_Name, Indicator_Code) %>% 
  unique()
le_indics
```
Life expectancy indicators: `r kable(le_indics)`


#### Find the average, minimum, and maximum female life expectancy at birth in 2016

```{r}
le_range <- 
  indicators %>% 
  filter(Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  summarise(Av_Life_Exp = mean(Yr2016, na.rm = TRUE),
            Min_Life_Exp = min(Yr2016, na.rm = TRUE),
            Max_Life_Exp = max(Yr2016, na.rm = TRUE))
le_range
```
The range of female life expectance, at birth in 2016, is: `r kable(le_range)`

#### Find the name of the country with the minimum female life expectancy at birth


```{r}
worst_le_coun <- 
  indicators %>% 
  filter(Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  filter(Yr2016 == min(Yr2016, na.rm = TRUE)) %>% 
  select(Country_Name, Yr2016)
worst_le_coun
```
The lowest life expectancy is in `r worst_le_coun[[1]]`.

#### Plot life expectancy at birth vs percentage of births attended by a qualified medic




```{r}
# Find out the name of the indicator and its code.
indicators %>% 
  filter(str_detect(Indicator_Name, "(?i)birth"),
         str_detect(Indicator_Name, "(?i)attend")) %>%
  select(Indicator_Name, Indicator_Code) %>% 
  unique()
```



```{r, include = TRUE}
# Make a plot.
plt.df <- indicators %>% 
  select(Country_Code, Indicator_Code, Yr2016) %>% 
  filter(Indicator_Code == "SH.STA.BRTC.ZS" | Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  filter(!is.na(Yr2016)) %>% 
  spread(key = Indicator_Code, value = Yr2016) %>% 
  drop_na() %>% 
  rename(Life_Exp = SP.DYN.LE00.FE.IN, Attended = SH.STA.BRTC.ZS)

ggplot(plt.df) +
  geom_point(aes(Attended, Life_Exp)) +
  geom_smooth(aes(Attended, Life_Exp), method = lm) +
  ggtitle("Life Expectancy vs Proportion of Attended Births")
```

Another version.

```{r, include = TRUE}
plt.df <- indicators %>% 
  select(Country_Code, Indicator_Code, Yr2016) %>% 
  filter(Indicator_Code == "SH.STA.BRTC.ZS" | Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  filter(!is.na(Yr2016)) %>% 
  spread(key = Indicator_Code, value = Yr2016) %>% 
  drop_na() %>% 
  rename(Life_Exp = SP.DYN.LE00.FE.IN, Attended = SH.STA.BRTC.ZS) %>% 
  left_join(country)

ggplot(plt.df) +
  geom_point(aes(Attended, Life_Exp, color = Income_Group, shape = Income_Group), size = 3) +
  geom_smooth(aes(Attended, Life_Exp), method = lm) +
  ggtitle("Life Expectancy vs Proportion of Attended Births")
```


#### Use a join to add Region and Income Group to the indicators table



```{r}
# Print the countries with life expectancy > 85 together with their region and Income group.
long_le_coun <- 
  indicators %>% 
  select(Country_Code, Country_Name, Indicator_Code, Yr2016) %>% 
  filter(Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  filter(Yr2016 > 85) %>% 
  rename(Fem_Life_Exp = Yr2016) %>%
  left_join(country) %>%
  select(Country_Name, Region, Fem_Life_Exp) %>% 
  arrange(desc(Fem_Life_Exp))
long_le_coun
```
Here are the longer lived countries: `r kable(long_le_coun)`



```{r}
# Print the countries with life expectancy < 60 together with their region and Income group.
short_le_coun <- 
  indicators %>% 
  select(Country_Code, Country_Name, Indicator_Code, Yr2016) %>% 
  filter(Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  filter(Yr2016 < 60) %>% 
  rename(Fem_Life_Exp = Yr2016) %>%
  left_join(country) %>%
  select(Country_Name, Region, Fem_Life_Exp) %>% 
  drop_na() %>% 
  arrange(desc(Fem_Life_Exp))
short_le_coun
```
And the short-lived countries: `r kable(short_le_coun)`

#### Calculate the average female life expectancy at birth in 2017 for each region and income group.



```{r}
# Rename the life expectancy column to something meaningful and order by region and life expectancy.
av_le <- 
  indicators %>% 
  select(Country_Code, Country_Name, Indicator_Code, Yr2016) %>% 
  filter(Indicator_Code == "SP.DYN.LE00.FE.IN") %>% 
  left_join(country) %>%
  select(Country_Name, Region, Income_Group, Yr2016) %>% 
  group_by(Income_Group, Region) %>% 
  summarise(Av_Fem_Life_Exp = mean(Yr2016, na.rm = TRUE)) %>% 
  drop_na() %>% 
  arrange(Region, desc(Av_Fem_Life_Exp))
av_le
```
Average life expectancies: `r kable(av_le)`

